Combinatorics of the geometry of Wilson loop diagrams

### 2019-02-06 General WLD combinatorics.tex
continuing slow progress on dimension lemmas

Let $\binom{[n]}{k}$ be the set of all $k$-subsets of the cyclically ordered set $[n]$.  For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\[] j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the \textit{Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we say that
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

\begin{dfn}\label{def:grassmann necklace}
A Grassmann necklace of type $(k,n)$ is a sequence $(I_1, \dots, I_n)$ of $n$ sets $I_i \in \binom{[n]}{k}$ such that for each $i \in [n]$:
\begin{itemize}
\item if $i \in I_i$, then $I_{i+1} = \big(I_i \backslash \{i\}\big) \cup \{j\}$ for some $j \in[n]$.
\item if $i \not\in I_i$, then $I_{i+1} = I_i$.
\end{itemize}
By convention, we set $I_{n+1} = I_1$.
\end{dfn}

@@

[Le diagrams]

@@

If $j \in [n]$ is a loop then $j \not\in I_j$, which can only happen if there are no propagators supported on vertex $j$. Conversely, if $j$ supports no propagators, then Algorithm ~\ref{alg:put GN on WLD} never assigns a propagator to $j$ and hence $j \not\in I_i$ for all $i \in [n]$.
\end{proof}

@@

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$.  Marcott in ***cite this*** has a different proof which is geometric and more elegant, bit it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

The dimension of a positroid cell is the same as the number of plusses in the associated Le diagram \cite[Theorem 6.5]{Postnikov}, and so with Algorithm~\ref{alg:put GN on WLD} along with the algorithm for passing from a Grassmann necklace to a Le diagram (reversing Oh's algorithm) described by some of us in \cite{reversingOh}, we can, at the cost of a bit of messy grunt work, give a recursive proof of the $3|\cP|$-dimensionality that explicitly describes the changes in the positions of the plusses in the Le diagram.

%In \cite{reversingOh}, Agarwala and Fryer give an algorithm for passing from a Grassmann Necklace to a Le diagram. We use this algorithm here to pass from a Wilson loop diagram to its associated Le diagram. In this manner, we show that the positroid cell defined by a Wilson loop diagram has dimension $3|\cP|$. Maybe say something about Amplituhedra having $4k$ dimension, but this is not quite what we have, since we are ignoring one column.
%
%When I speak of a \emph{WLD}, I mean one which satisfies your density hypothesis and other standard hypotheses.  When I speak of a propagator \emph{contributing} a vertex to an element of the Gra\ss mann\footnote{I suppose it is a bit of an affectation to use an eszett when writing in English, and actually the eszett is kind of ugly in this font, but since I've started to I'll stick with it for now but you certainly shouldn't feel obliged to do it too.} necklace I mean that according to the rule you guys have to build the Gra\ss mann necklace from the diagram by starting at a vertex and taking the clockwise-most covering propagator which hasn't already been taken, when a propagator is taken then it is contributing that vertex.  Also, I will be using your algorithm to convert the Gra\ss mann necklace into a Le diagram by non-intersecting paths.
%
%
%We need four lemmas from stuff you guys have already figured out.  The first one is a corollary of your lemma characterizing intervals contributed by each covered vertex of a
%propagator.
%
%The first lemma is Lemma~\ref{vertex cyclic int lem} and I moved it up to a previous section.
%
%The second lemma is the fact that we understand uncovered vertices.  You probably see many ways to prove this from things you already know.  One would be to say that from your algorithm the column of the uncovered vertex simply plays no role.

We need a number of lemmas, of roughly increasing degree of technicality.

\begin{lem}\label{lem uncovered}
  Let $W$ be an admissible Wilson loop diagram with $k$ propagators and with a vertex $i$ which supports no propagators.  Let $V$ be $W$ with vertex $i$ removed.  Then the Le diagram of $W$ is the Le diagram of $V$ with an extra column of all $0$s inserted so as to be labelled $i$ on its bottom in the Oh labelling of the border of the shape of the Le diagram
\end{lem}

\note{I'm not sure how much I should assume here in terms of what the reader already knows}

@@

By the algorithm of \cite{reversingOh} [ref][Sian: update this when/if you put this alg in section 3.1], since $i$ never appears in the Grassmann necklace of $W$, the column labelled $i$ is filled with all $0$s.  Otherwise the algorithm of \cite{reversingOh} behaves identically on $V$ and $W$ with the plusses for the incremented indices shifted one column to the left, thus skipping over column $i$ in the Le diagram for $W$.
\end{proof}

%The third lemma is now lemma~\ref{lem sian}.

%The fourth lemma is that we can rotate or reflect without changing the number of plusses.  The best way to see this is probably geometrically so I leave that up to you.

@@

%\subsubsection{Changes in Gra\ss mann necklaces for nice configurations}

%Given a propagator $p$ in a WLD $D$, note that $p$ divides remaining propagators of $D$ into two sets depending on which side of $p$ they live.

@@

%Given an admissible Wilson loop diagram $W$, $I_i^{D}$ denotes the $i$th Grass mann necklace element corresponding to $W$.

Our goal is to compute the dimension of the positroid associated to an arbitrary admissible Wilson loop diagram. By Lemmas \ref{lem uncovered}, \ref{lem dihedral}, and \ref{lem good p}, it suffices to study the Le diagrams of \hlfix{weakly}{check} admissible Wilson loop diagrams admitting one of the configurations described in Lemma~\ref{lem good p} with propagators $p$ and $q$ (if it exists) both of length 2. Diagrams of this type are the subject of the next few lemmas, and are illustrated in Figure~\ref{fig special p}.

\begin{lem}\label{lem I}
Let $W$ be an admissible Wilson loop diagram with $n\geq 1$ propagators, and suppose that $W$ admits one of the configurations described in Lemma~\ref{lem good p} with $p$ and $q$ (if it exists) both of length 2. Let $V$ be $W$ with $p$ removed.  Then
  \begin{align*}
    I_1^{W} & = I_1^{V} \cup \{n-3\} \\
    I_n^{W} & = I_1^{V} \cup \{n\} \\
    I_{n-1}^{W} & = I_n^{V} \cup \{n-1\} \\
    I_{n-2}^{W} & =
    \begin{cases}
      I_{n-2}^{V}\cup \{n-2\} & \text{if $n-2\not\in I_{n-2}^{V}$} \\
      I_{n-2}^{V}\cup \{n-1\} & \text{if $n-2\in I_{n-2}^{V}$, $n-1\not\in I_{n-2}^{V}$} \\
      (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\} & \text{if $n-1, n-2\in I_{n-2}^{V}$}
    \end{cases} \\
    I_{k}^{W} & =
    \begin{cases}
      I_k^{V}\cup \{n-3\} & \text{if $n-3 \not\in I_k^{V}$}\\
      I_k^{V}\cup\{n-2\} & \text{if $n-3\in I_k^{V}$}
    \end{cases} \\
    & \text{for $1<k<n-2$}
  \end{align*}
\end{lem}

@@

Next, consider $I_n^{(W)}$, i.e. we apply the Grassmann necklace algorithm with starting vertex $n$. In both cases$I_n^{(W)}(p) = n$. We are now at vertex 1, and the unassigned propagators are exactly those that appear in $V$. Therefore the algorithm continues as in $I_1^{(V)}$, i.e. we have $I_n^{W} = I_1^{V} \cup \{n\}$. Similarly, we must have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$.

@@

\begin{lem}\label{lem shape}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}.
  The shape of the Le diagram of $V$ can be built from left to right of the following blocks: a rectangle with 3 columns, one more column of the same length, a partition shape with at most as many rows as the rectangle.
  The shape of the Le diagram of $W$ can be built from left to right of the following blocks: a rectangle with 3 columns and one more row than the first rectangle of $V$, the same partition shape as in $V$.
\end{lem}

\begin{proof}
  $I_1$ determines the shape of the Le diagram.
  By Lemma~\ref{lem I}, $I_1^{W} = I_1^{V}\cup \{n-3\}$.  This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, a new column fourth from the left.
\end{proof}

The shapes of the Le diagrams of $V$ and $W$ are illustrated in Figure~\ref{fig Le}.  The pieces of the Le diagrams will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows, as in the figure.  Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of the Le diagrams of $V$ and $W$ are identical and the plusses in the $\mathcal{A}$ parts are very closely related.  When we speak of a plus in the Le diagram of $V$ being the same as in $W$ or vice versa, we mean that the plus' position in $\mathcal{A}$ or $\mathcal{B}$ is the same.  Because of the column insertion the absolute indices may differ.

@@

Suppose we are following the Grassmann necklace to Le diagram algorithm, and we put a plus in a box because of a path from vertical boundary edge $i$ to bottom boundary edge $j$.  Then say this plus is in the $i\rightarrow j$ position.

\begin{lem}\label{lem n and n-1}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}.
  Then $I_n^{W}$ and $I_{n-1}^{W}$ give all the same plusses as $I_n^{V}$ along with plusses in the leftmost two boxes of the bottom row of the Le diagram of $W$.
\end{lem}

\begin{proof}
  By Lemma~\ref{lem I} $I_n^{W}= I_1^{V} \cup \{n\}$, so by the Grassmann necklace to Le diagram algorithm the only plus this builds in the Le diagram of $W$ is  the one in the $n-3\rightarrow n$ position, that is in the leftmost box of the bottom row.

  Also by Lemma~\ref{lem I} $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$.  Additionally $n-3\not\in I_{n}^{V}$ since if it were then $n-3$ would also be in $I_1^{V}$ and hence propagator $p$ could not contribute $n-3$ to $I_1^{W}$ in contradiction to Lemma~\ref{vertex cyclic int lem}.  Similarly $n-1, n-2\not\in I_n^{V}$.  Thus the paths putting the plusses in from $I_n^{V}$ lie completely in $\mathcal{B}$ or take some vertical boundary edge $>n-3$ to $n$.   Now view these paths in the Le diagram of $W$ and note that the path $n-3\rightarrow n-1$ is compatible, and so these paths together build the plusses that $I_{n-1}^{W}$ contributes.  That is, we get all the plusses from $I_{n}^{V}$ along with a $n-3\rightarrow n-1$ plus, that is a plus in the second to the right box of the bottom row.
\end{proof}

\begin{lem}\label{lem n-2 good}
  Let $V$ and $W$ be as in Lemma~\ref{lem I} with $n-2 \not\in I_{n-2}^{V}$.

  Then $I_{n-2}^{W}$ gives
  an $n-3\rightarrow n-2$ plus and
  all the $I_{n-1}^{V}=I_{n-2}^{V}$ plusses.
\end{lem}

\begin{proof}
  If $n-2\not\in I_{n-2}^{V}$ then $I_{n-1}^{V}=I_{n-2}^{V}$ and by Lemma~\ref{lem I} $I_{n-2}^{W} = I_{n-1}^{V} \cup \{n-2\}$.  Note that $n-3\not\in I_{n-2}^{V}$ by Lemma~\ref{vertex cyclic int lem}.  Therefore the paths for $I_{n-2}^{W}$ are the paths for $I_{n-1}^{V}$ along with the $n-3\rightarrow n-2$ path.  This gives the statement of the lemma.

@@

\begin{lem}\label{lem n-2 bad}
  Let $V$ and $W$ be as in Lemma~\ref{lem I} with $n-2,n-1 \in I_{n-2}^{V}$.

  Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ give the following plusses:
  \begin{itemize}
  \item An $n-3\rightarrow n-2$ plus and an $n-5\rightarrow n-1$ plus.
  \item All the $I_{n-1}^{V}$ plusses.
  \item $I_{n-2}^{V}$ gives an $n-5\rightarrow n-2$ plus and no other term in the Gramann necklace of $V$ gives a plus in this column.  This $+$ does not appear in $W$ from $I_{n-2}^{W}$ but an $n-5\rightarrow n-1$ plus does instead.
  \item All other plusses of $I_{n-2}^{V}$.
  \item $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted over into the $n-2$ column in $W$.
  \item All other plusses of $I_{n-3}^{V}$.
  \end{itemize}
  Furthermore, no element of the Gramann necklace of $V$ gives an $n-5\rightarrow n-1$ plus.
\end{lem}

>>>>

Let $\binom{[n]}{k}$ be the set of all $k$-subsets of the cyclically ordered set $[n]$.  For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

\begin{dfn}\label{def:grassmann necklace}
A {\bf Grassmann necklace} of type $(k,n)$ is a sequence $(I_1, \dots, I_n)$ of $n$ sets $I_i \in \binom{[n]}{k}$ such that for each $i \in [n]$:
\begin{itemize}
\item if $i \in I_i$, then $I_{i+1} = \big(I_i \backslash \{i\}\big) \cup \{j\}$ for some $j \in[n]$.
\item if $i \not\in I_i$, then $I_{i+1} = I_i$.
\end{itemize}
By convention, we set $I_{n+1} = I_1$.
\end{dfn}

@@

\begin{dfn}\label{def:le diagram}
A {\bf Le diagram} is a Young diagram in which every square contains either a $+$ or a $0$, subject to the rule that if a square contains a $0$ then either all squares to its left (in the same row) must also contain a $0$, or all squares above it (in the same column) must also contain a $0$, or both.
\end{dfn}

By [ref], the set of all Le diagrams that fit within a $k\times(n-k)$ rectangle is in 1-1 correspondence with the positroid cells of $\Gr(k,n)$. The dimension of a positroid cell is equal to the number of $+$ squares in its Le diagram [ref].

The rows and columns of a Le diagram are labelled as follows: given a Le diagram fitting inside a $k\times (n-k)$ box, arrange the numbers $1,2, \dots, n$ along its southeast border, starting from the top-right corner. See Figure \ref{fig:row column numbering} for examples.

\begin{figure}[h!]
\[\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (5,0) -- (5,2);

\node at (5.15,2.5) {\footnotesize$1$};
\node at (4.5,1.85) {\footnotesize$2$};
\node at (4.15,1.5) {\footnotesize$3$};
\node at (3.5,0.85) {\footnotesize$4$};
\node at (2.5,0.85) {\footnotesize$5$};
\node at (2.15,0.5) {\footnotesize$6$};
\node at (1.5,-0.15) {\footnotesize$7$};
%\node at (0.5,-0.15) {\footnotesize$8$};
\end{tikzpicture}
\qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (6,0) -- (6,3) -- (5,3);

\node at (5.5,2.85) {\footnotesize$1$};
\node at (5.15,2.5) {\footnotesize$2$};
\node at (4.5,1.85) {\footnotesize$3$};
\node at (4.15,1.5) {\footnotesize$4$};
\node at (3.5,0.85) {\footnotesize$5$};
\node at (2.5,0.85) {\footnotesize$6$};
\node at (2.15,0.5) {\footnotesize$7$};
\node at (1.5,-0.15) {\footnotesize$8$};
%\node at (0.5,-0.15) {\footnotesize$9$};
\end{tikzpicture}
\]
\caption{Row and column numbering for a Young diagram with $k = 3$, $n = 7$ (left) and $k = 3$, $n = 8$ (right). The top left box in each diagram has coordinates $(1,7)$ (left diagram), $(2,8)$ (right diagram).}
\label{fig:row column numbering}
\end{figure}

An algorithm for constructing the Le diagram associated to a Grassmann necklace was given by Agarwala and Fryer in [ref]. Since we will make use of this algorithm in Section [ref] below, we summarise the process here.
\begin{algorithm}\label{alg:GN to Le} \ [ref]
Let $(I_1,\dots,I_n)$ be a Grassmann necklace of type $(k,n)$. Within a $k \times(n-k)$ square, draw the Young diagram whose rows are labelled by $I_1$ (as per the convention above).

For each $i$, $2 \leq i \leq n$:
\begin{itemize}
\item Write \[I_1 \setminus I_i = \{a_1 > a_2 > \dots > a_r\}, \quad I_i \setminus I_1 = \{b_1 < b_2 < \dots < b_r\},\]
where the inequalities denote the $<_1$ order (subscripts suppressed for clarity).
\item For $1 \leq j \leq r$ , place a $+$ in square $(a_j,b_j)$ of the diagram. (We will sometimes refer to this $+$ as being {\em in the $a_j \rightarrow b_j$ position}.)
\end{itemize}
After performing the above for $2 \leq i \leq n$, place a $0$ in any remaining unfilled boxes.
\end{algorithm}

An algorithm for constructing the Grassmann necklace of a Le diagram also exists; this was given by Oh in [ref]. A method for using the Le diagram to test whether a given $k$-subset is a basis for the corresponding positroid or not was given by Casteels in [ref].

@@

If $j \in [n]$ is a loop then $j \not\in I_j$, which can only happen if there are no propagators supported on vertex $j$. Conversely, if $j$ supports no propagators, then Algorithm~\ref{alg:put GN on WLD} never assigns a propagator to $j$ and hence $j \not\in I_i$ for all $i \in [n]$.
\end{proof}

@@

Our next goal is to show that the dimension of the positroid cell defined by a Wilson loop diagram $(\cP, [n])$ has dimension $3|\cP|$.  Marcott in ***cite this*** has a different proof which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  Our approach is much more explicit.

Recall that the dimension of a positroid cell is equal to the number of plusses in the associated Le diagram \cite[Theorem 6.5]{Postnikov}. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from WLD to Grassmann necklace to Le diagram) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams, and hence give a recursive proof of the $3|\cP|$-dimensionality of the cells.

We start with several lemmas, of roughly increasing degree of technicality.

\begin{lem}\label{lem uncovered}
Let $W$ be an admissible Wilson loop diagram with $k$ propagators, and with a vertex $i$ which supports no propagators.  Let $V$ be $W$ with vertex $i$ removed.  Then the Le diagram of $W$ is obtained from the Le diagram of $V$ by inserting an extra column containing all $0$s in position $i$ (i.e. such that the new column has the label $i$).
\end{lem}

@@

By Algorithm~\ref{alg:GN to Le}, since $i$ never appears in the Grassmann necklace of $W$ we know that the column labelled $i$ contains only $0$ boxes.  Otherwise the algorithm of \cite{reversingOh} behaves identically on $V$ and $W$ with the plusses for the incremented indices shifted one column to the left, thus skipping over column $i$ in the Le diagram for $W$.
\end{proof}

@@

@@

Combining Lemmas \ref{lem uncovered}, \ref{lem dihedral}, and \ref{lem good p}, it therefore suffices to study the Le diagrams of \hlfix{weakly}{check} admissible Wilson loop diagrams admitting one of the configurations described in Lemma~\ref{lem good p} with propagators $p$ and $q$ (if $q$ exists) both of length 2. See Figure~\ref{fig special p} for an illustration of the two possibilities.

The next few lemmas describe how diagrams of this type are related to the corresponding diagram with propagator $p$ removed, first in terms of the Grassmann necklaces and then in terms of the Le diagrams. These technical lemmas will form the backbone of the inductive step in the main dimensionality argument.

\begin{lem}\label{lem I}
Let $W$ be an admissible Wilson loop diagram with $n\geq 1$ propagators, and suppose that $W$ admits one of the configurations described in Lemma~\ref{lem good p}, with $p$ and $q$ (if $q$ exists) both of length 2. Let $V$ be $W$ with $p$ removed.  Then
  \begin{align*}
    I_1^{W} & = I_1^{V} \cup \{n-3\} \\
    I_n^{W} & = I_1^{V} \cup \{n\} \\
    I_{n-1}^{W} & = I_n^{V} \cup \{n-1\} \\
    I_{n-2}^{W} & =
    \begin{cases}
      I_{n-2}^{V}\cup \{n-2\} & \text{if $n-2\not\in I_{n-2}^{V}$} \\
      I_{n-2}^{V}\cup \{n-1\} & \text{if $n-2\in I_{n-2}^{V}$, $n-1\not\in I_{n-2}^{V}$} \\
      (I_{n}^{V} - \{n-5\})\cup \{n-1,n-2\} & \text{if $n-1, n-2\in I_{n-2}^{V}$}
    \end{cases} \\
    I_{k}^{W} & =
    \begin{cases}
      I_k^{V}\cup \{n-3\} & \text{if $n-3 \not\in I_k^{V}$}\\
      I_k^{V}\cup\{n-2\} & \text{if $n-3\in I_k^{V}$}
    \end{cases} \\
    & \text{for $1<k<n-2$}
  \end{align*}
\end{lem}

@@

Next, consider $I_n^{(W)}$, i.e. we apply the Grassmann necklace algorithm with starting vertex $n$. In both cases we have $I_n^{(W)}(p) = n$. We are now at vertex 1, and the unassigned propagators are exactly those that appear in $V$. Therefore the algorithm continues as in $I_1^{(V)}$, i.e. we have $I_n^{W} = I_1^{V} \cup \{n\}$. Similarly, we must have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$.

@@

\begin{lem}\label{lem shape}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}.
  The shape of the Le diagram of $V$ can be built from left to right of the following blocks: a rectangle with 3 columns, one more column of the same length, a partition shape with at most as many rows as the rectangle.
  The shape of the Le diagram of $W$ can be built from left to right of the following blocks: a rectangle with 3 columns and one more row than the first rectangle of $V$, and the same partition shape as in $V$.
\end{lem}

\begin{proof}
See Figure~\ref{fig Le} for an illustration of the shapes described in the statement of the lemma.

Recall that $I_1$ determines the shape of the Le diagram. By Lemma~\ref{lem I}, we know that $I_1^{W} = I_1^{V}\cup \{n-3\}$, and that $n,n-1,n-2, n-3 \not\in I_1^{(V)}$. This implies that the right hand boundary of the shape of $V$ is the same as the right hand boundary of the shape of $W$ except that $W$ has one additional row of 3 boxes while $V$ has an additional column in the $n-3$ position; that is, an extra column fourth from the left.
\end{proof}

The pieces of the Le diagrams of $V$ and $W$ will be called $\mathcal{A}$ and $\mathcal{B}$ in what follows, as in Figure~\ref{fig Le}. Over the course of the next few lemmas we will prove that the plusses in the $\mathcal{B}$ parts of each diagram are identical, and that the plusses in the $\mathcal{A}$ parts are very closely related.

Note that when we speak of a plus in the Le diagram of $V$ being the same as in $W$ or vice versa, we mean that the position of the plus in $\mathcal{A}$ or $\mathcal{B}$ is the same; because of the column insertion the absolute indices may differ.

@@

%Recall Algorithm~\ref{alg:GN to Le}, which constructs the Le diagram associated to a Grassmann necklace. As described in Section [ref], if Algorithm~\ref{alg:GN to Le} places a $+$ in the box with row index $i$ and column index $j$, we say that this plus is in the $i \rightarrow j$ position.

\begin{lem}\label{lem n and n-1}
Let $V$ and $W$ be as in Lemma~\ref{lem I}.
Then $I_n^{W}$ and $I_{n-1}^{W}$ together yield the same plusses as $I_n^{V}$ did, along with extra plusses in the leftmost two boxes of the bottom row of the Le diagram of $W$.
\end{lem}

\begin{proof}
By Lemma~\ref{lem I} we have $I_n^{W}= I_1^{V} \cup \{n\}$ and $I_1^{(W)} = I_1^{(V)} \cup \{n-3\}$. Thus 
\[I_1^{(W)} \setminus I_n^{(W)} = \{n-3\}, \qquad \qquad I_n^{(W)} \setminus I_1^{(W)} = \{n\},\]
and so by Algorithm~\ref{alg:GN to Le} we have a plus in the $(n-3) \rightarrow n$ position, i.e. in the leftmost box of the bottom row.

Also by Lemma~\ref{lem I} we have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$. Recall from the proof of Lemma~\ref{lem I} that $n-3,n-2,n-1,n \not\in I_1^V$; therefore $I_1^W\setminus I_{n-1}^W = (I_1^V \setminus I_n^V) \cup\{n-3\}$, and $n-3$ is maximal in this set. Similarly, 
\[I_{n-1}^W \setminus I_1^W = (I_n^V \setminus I_1^V) \cup \{n-1\} \subseteq \{n-1,n\},\]
where the final inclusion follows from the definition of Grassmann necklace. In particular, $n-1$ is minimal in this set, so Algorithm~\ref{alg:GN to Le} yields a plus in the $(n-3) \rightarrow (n-1)$ position (i.e. the second box in the bottom row of the Le diagram of $W$) along with any plusses yielded by $I_n^V$.

% Additionally $n-3\not\in I_{n}^{V}$ since if it were then $n-3$ would also be in $I_1^{V}$ and hence propagator $p$ could not contribute $n-3$ to $I_1^{W}$ in contradiction to Lemma~\ref{vertex cyclic int lem}.  Similarly $n-1, n-2\not\in I_n^{V}$.  Thus the paths putting the plusses in from $I_n^{V}$ lie completely in $\mathcal{B}$ or take some vertical boundary edge $>n-3$ to $n$.   Now view these paths in the Le diagram of $W$ and note that the path $n-3\rightarrow n-1$ is compatible, and so these paths together build the plusses that $I_{n-1}^{W}$ contributes.  That is, we get all the plusses from $I_{n}^{V}$ along with a $n-3\rightarrow n-1$ plus, that is a plus in the second to the right box of the bottom row.
\end{proof}

\begin{lem}\label{lem n-2 good}
  Let $V$ and $W$ be as in Lemma~\ref{lem I}, with $n-2 \not\in I_{n-2}^{V}$. Then $I_{n-2}^{W}$ contributes an $(n-3)\rightarrow (n-2)$ plus, and all of the $I_{n-1}^{V}=I_{n-2}^{V}$ plusses.
\end{lem}

\begin{proof}
If $n-2\not\in I_{n-2}^{V}$ then $I_{n-1}^{V}=I_{n-2}^{V}$ and by Lemma~\ref{lem I} we have $I_{n-2}^{W} = I_{n-1}^{V} \cup \{n-2\}$.  Note that $n-3\not\in I_{n-2}^{V}$ by Lemma~\ref{lem no fourth vertex}.  Therefore the paths for $I_{n-2}^{W}$ are the paths for $I_{n-1}^{V}$ along with the $(n-3)\rightarrow (n-2)$ path.  This gives the statement of the lemma.

@@

\begin{lem}\label{lem n-2 bad}
Let $V$ and $W$ be as in Lemma~\ref{lem I}, and suppose that $n-2,n-1 \in I_{n-2}^{V}$.

Then $I_{n-2}^{W}$ and $I_{n-3}^{W}$ contribute the following plusses to the Le diagram of $W$
  \begin{itemize}
  \item An $(n-3)\rightarrow (n-2)$ plus and an $(n-5)\rightarrow (n-1)$ plus.
  \item All of the $I_{n-1}^{V}$ plusses.
  \item In the Le diagram of $V$, $I_{n-2}^{V}$ contributes an $n-5\rightarrow n-2$ plus and no other term in the Grassmann necklace of $V$ gives a plus in this column.  $I_{n-2}^W$ does not contribute this $+$ but yields an $n-5\rightarrow n-1$ plus instead.
  \item All other plusses of $I_{n-2}^{V}$.
  \item In the Le diagram of $V$, $I_{n-3}^{V}$ gives a plus in the $n-3$ column.  This $+$ is shifted over into the $n-2$ column in $W$.
  \item All other plusses of $I_{n-3}^{V}$.
  \end{itemize}
Furthermore, no element of the Gramann necklace of $V$ gives an $n-5\rightarrow n-1$ plus.
\end{lem}
